\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-04-07 abstract.tex
.

In this work we address the problem of finding unsafe behaviors with
given a safety property for a hybrid dynamical system. We treat the
system as a black box and approach the problem in two distinct steps.
First, we model the behavior of the system using a piecewise affine
discrete time model. Such a model is incomplete in nature and is
computed with respect to the given property. Next, we encode the
search for falsification as a bounded model checking query and use an
SMT solver to find a counter example. If found, we check the existance
of the violation in the original system to ensure reproducibility.

%Time bounded LTL property in 
%Models from verificaiton
%model without data
%black box model/data -> structured model

>>>>

Sampling based falsification techniques for black box hybrid dynamical
systems use numerical simulations to search for violations of
properties (specified in temporal logic). This constraints them to a
`best effort' category unlike rigorous verification techniques, which
however require a white box model. This work tries to bridge the gap
by presenting an automatic falsification technique which models the
black box system before searching for a violation. It proceeds by (a)
modeling the behavior of the system using a piecewise affine (PWA)
discrete time model and, (b) encoding the search for a violation as a
standard bounded model checking problem. The resulting problem is a
disjunction of linear programs and can be solved directly using
off-the-shelf SMT solvers or by multiple runs of a linear programming
solver. If a counter-example is found, the existence of a
corresponding violation in the original system is checked for
reproducibility.

% In this work we address the problem of finding unsafe behaviors with
% given a safety property for a hybrid dynamical system. We treat the
% system as a black box and approach the problem in two distinct steps.
% First, we model the behavior of the system using a piecewise affine
% discrete time model. Such a model is incomplete in nature and is
% computed with respect to the given property. Next, we encode the
% search for falsification as a bounded model checking query and use an
% SMT solver to find a counter example. If found, we check the existence
% of the violation in the original system to ensure reproducibility.

%Time bounded LTL property in
%Models from verification
%model without data
%black box model/data -> structured model

### 2017-04-07 bmc.tex
.

% include an algorithm
We now describe the entire procedure in three steps.
\begin{enumerate}
    \item Given a black box system $\System$ specified by
        $\simulate_\System$, we
        use scatter-and-simulate with ($\Delta$-length) trajectory
        segments to sample a finite reachability graph
        $\scr{H}(\Delta)$.
    \item Using regression, we enrich the graph relations and annotate
        the edges with the estimated affine maps.
    \item Interpreting the directed reachability graph as a transition
        system, we use a bounded model checker to find fixed length
        violations to safety properties.
\end{enumerate}

@@

We now describe the above using the Van der Pol oscillator benchmark
presented in \secref{math-dyn-sys}. Simulations generated using
uniformly random samples are shown in \figref{vdp-cont}. We want to
check the property $P_3$, indicated by the red box, given the initial
set indicated by the green box. The abstraction is defined by
$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
where each cell is of size $0.2 \times 0.2$ units.
Scatter-and-simulate is then used to construct the abstract graph $G$,
using $2$ samples per cell and the time step $\Delta = 0.1s$. The
complete process follows.

@@

\begin{table*}[!htbp]
\centering
\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
\label{tab:vdp-pwa}
\begin{tabular}{@{}ccccc@{}}
\toprule
$C$ & $C'$ & $A$ & $b$ & $\delta$\\
\midrule
(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
\midrule
(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
\midrule
(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
\bottomrule
\end{tabular}
\end{table*}

>>>>

% include an algorithm
We now describe the entire procedure in three steps.
\begin{enumerate}
    \item Given a black box system $\System$ specified by
        $\simulate^\System$, we
        use scatter-and-simulate with ($\Delta$-length) trajectory
        segments to sample a finite reachability graph
        $\scr{H}(\Delta)$.
    \item Using regression, we enrich the graph relations and annotate
        the edges with the estimated affine maps.
    \item Interpreting the directed reachability graph as a transition
        system, we use a bounded model checker to find fixed length
        violations to safety properties.
\end{enumerate}

@@

We now describe the above using the Van der Pol oscillator benchmark
presented in \cite{zutshi2014multiple}. Simulations generated using
uniformly random samples are shown in \figref{vdp-cont}. We want to
check the property $P_3$, indicated by the red box, given the initial
set indicated by the green box. The abstraction is defined by
$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
where each cell is of size $0.2 \times 0.2$ units.
Scatter-and-simulate is then used to construct the abstract graph $G$,
using $2$ samples per cell and the time step $\Delta = 0.1s$. The
complete process follows.

@@

\begin{table}[!htbp]
\centering
\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
\label{tab:vdp-pwa}
\begin{tabular}{@{}ccccc@{}}
\toprule
$C$ & $C'$ & $A$ & $b$ & $\delta$\\
\midrule
(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
\midrule
(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
\midrule
(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
\bottomrule
\end{tabular}
\end{table}

### 2017-04-07 implementation.tex
.

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
Benchmark & BMC & S3CAM-R & S3CAM\\
\midrule
Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
Brusselator               &$<$1  & $4$    & $2$\\
Lorenz                    &$<$1 & $122$  & $35$\\
%B.Ball                    &  &   & \\
Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
Bouncing Ball             &-  &$TO$   & $450$\\

>>>>

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}lllll@{}}
\toprule
    Benchmark & BMC & LP & S3CAM-R & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$<$1 & $0$ & $130$ & $24$\\
    Brusselator               &$<$1 & $0$ & $4$    & $2$\\
    Lorenz                    &$<$1 & $0$ & $122$  & $35$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$4480$ & $0$ & $9423$   &$603$ \\
    Nav ($\prop_Q$)           &$970$ & $0$ & $8105$   &$546$ \\
    Nav ($\prop_R$)           &-  & $0$ &$TO$ & $2003$\\
    Nav ($\prop_S$)           &-  & $0$ &$TO$ & $2100$\\
    Bouncing Ball             &-  & $0$ &$TO$ & $450$\\

### 2017-04-07 prelims.tex
.

In this section, we present the basics of relational abstractions and
fundamentals of regression analysis.

\subsection{Relational Abstraction}

For general hybrid dynamical systems, algorithmic ways for computing
analytical solutions do not exist. Discretization (along with suitable
abstractions) is often employed to transform the systems into a
discrete transition system. Relational abstractions is one such idea,
which abstracts continuous dynamics by discrete relations using
appropriate \textit{reachability invariants}. Both time independent
and time dependent relations have been proposed; the former captures
all reachable states over all time, whereas, the latter explicitly
includes time by relating reachable states to time. Thus it
can prove timing properties whereas, time independent relational
abstractions can not. In both cases, the resulting abstractions can be
interpreted as discrete transition systems and can be analyzed using
model checkers. We briefly discuss these two types of abstractions.

\mypara{Timeless Relational Abstraction}

Relational abstractions for hybrid systems were proposed in
\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
model, they can summarize the continuous dynamics of each mode using a
binary relation over continuous states. The resulting relations are
timeless (independent) and hence valid for all time as long as
the mode invariant is satisfied. The relations take the general form
of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
relational operators $=, \ge, \le, <, >$. For example, an abstraction
that captures the monotonicity with respect to time for the
differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
capturing the relation between the set of ODEs: $\dot{x} = 2$,
$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.

Such relation summaries, discretize the continuous evolution of a
given system into an infinite state transition system, in which safety
properties can be verified using $k$-induction, or falsified using
bounded model checkers. The relationalization procedure involves
finding suitable invariants in a chosen abstract domain, such as
affine abstractions, eigen abstractions or box abstractions
\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
techniques like template based invariant generation
\cite{Gulwani+Tiwari/2008/Constraint,
Colon+Sankaranarayanan+Sipma/03/Linear} can be used.

\mypara{Timed and Time-Aware Relational Abstractions}

As the above discussed relations are timeless, we cannot reason over
the timing properties of the original system. Moreover, they are not
suitable for time-triggered systems. Timed relational
abstractions~\cite{zutshi2012timed} and time-aware relational
abstractions~\cite{mover2013time} were proposed to overcome these
shortcomings by explicitly including time in the relations. In
addition, both can be more precise than their timeless counterpart,
but unlike it, assume continuous affine dynamics.

To analyze time-triggered systems like an SDCS modeled by an affine
hybrid automata~\footnote{An affine hybrid automata is restricted to
the ODEs, resets and guards being affine.}, timed relational
abstractions can be computed directly from the solutions of the
underlying affine ODEs.  Recall that the solution of an affine ODE
$\dot{\x} = A\x$ can be represented using the matrix exponential as
$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
However, for the case of SDCS with a fixed sampling time period, we
obtain linear relations in $\x$.  We demonstrated the usefulness of
timed relational abstractions for the case of linear systems
in \cite{zutshi2012timed}, and now incorporate the idea to discretize
black box dynamical systems.

Similar to timeless relational abstractions, time-aware relational
abstractions \cite{mover2013time} construct binary relations between
the current state $\x$ of the system and any future reachable state
$\x'$. The difference lies with the latter also constructing relations
between the current time $t$ and any future time $t'$. This is
achieved by a case by case analysis of the eigen structure of the
matrix $A$ (of an affine ODE). Separate abstractions are used for the
case of linear systems with constant rate, real eigen values and complex eigenvalues.

@@

\mypara{Regression Analysis}

In statistics, regression is the problem of finding a
\textit{predictor}, which can suitably predict the relationship
between the given set of observed input $\x$ and output $\y$ vectors.
In other words, assuming that $\y$ depends on $\x$, regression
strategies find either a parameterized or a non-parameterized
prediction function to explain the dependence. We now discuss simple
linear regression, which is parametric in nature and searches for an
affine predictor. It is also called \textit{ordinary least squares}.

@@

Let the data set be comprised of $N$ input and output pairs $(\x,
y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
the case in the current context of \textit{finding the best fit}, the
problem is over-determined; there are more equations than
unknowns. Hence, a single affine function cannot be found which
satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
need to find the `best' choice for $\vec{a}$ and $b$. This is formally
defined using a loss function. For the case of simple linear
regression or OLS, the loss function is the sum of squares of the
errors in prediction.  The task is then to determine the vector of
coefficients $\vec{a}$ and an offset constant $b$, such that the least
square error of the affine predictor is minimized for the given data
set.
\begin{equation}
    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
\label{reg}
\end{equation}
To ease the presentation, we can rewrite the above as a homogeneous
expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
\[
    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
\]
The solution of OLS can be analytically computed as
\[ Ab = (X^TX)^{-1}X^T\y\]
where $X$ is the matrix representing the horizontal stacking of all
$\hat{\x}$.
The details can be found in several texts on learning and statistics
\cite{friedman2001elements}.

@@

@@

We can now use a PWA transition system $\rho$ to approximate the
behavior of a hybrid dynamical system $\System$ defined by a simulator
$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
form, we can only approximate general hybrid systems.}. Each
transition of $\rho$ represents a discrete time step $\Delta$. Let the
state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
predicates $g_\tau$ are given by a conjunction of hyper-planes and
hence can be represented in matrix form as a polyhedron (defined by
$m$ constraints) in the state-space
\[
g:C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
$m$. The affine maps $f_\tau$ can be represented in matrix form as
$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
an error term $\delta$ which is defined by a vector of intervals
$[\delta_{\min},\delta_{\max}]$. Hence,
\[
f:\x \mapsto A\x + \vb + \delta
\]
defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
\in f_\tau(\x)}$.

@@

>>>>

In this section, we present the model of the hybrid dynamical system
under test and the basics of pwa relational abstraction.

We now use the family of reachability relations presented
in~\cite{zutshi2014multiple} to charecterize the system $\System$
under test. We denote the states of the hybrid system $\System$, by $\HybridStates \subseteq \Modes \times \Reals^n$ be the (infinite)
set of hybrid states of the system $\System$, where $\Modes$ is a
finite set of discrete modes, and $\Inputs$ be the set of input
signals to $\System$ of the form $[0,T] \rightarrow \reals^k$ for some
given time horizon $T$. The family of
relations $\reach{t,u} \subseteq \HybridStates \times \HybridStates$,
parameterized by time $t \geq 0$ and input signal $u \in \Inputs$
defined over time $[0,T]$.  We use the notation $\vx\ \reach{t,u} \vy$
to denote $(\vx,\vy) \in \reach{t,u}$. The relations $\reach{t,u}$
satisfy the following basic properties:

We assume a black-box model
equipped with a forward simulation map $\simulate^{\System}$ that can
be used to compute the discrete time traces of the system. Let
$\reach{}$ Given the
hybrid state space $\vx\in\HybridStates$ and input space $u
\in\Inputs$ and time step $\Delta_t$, it computes the unique successor
state $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u}
\vy$.

\begin{definition}

    The system under test $\System$ evolves in the set of infinite
    hybrid state-space $\HybridStates$ under the effect of inputs $\vu
    \in \Inputs$. Its evolution is defined by a simulation function
    $\simulate^\System : \HybridStates \times \Inputs \Reals^+ \mapsto
    \HybridStates$, where $\simulate^\System(\x,\vu,\Delta_t)$ maps a
    current state $\x$ at time $t$ to a new state $\x'$ at time
    $t+\Delta_t$, under a constant input $\vu$.

The subscript $\System$ will be dropped whenever the system $\System$
being referred to is clear from the context.

\myipara{Black-Box Assumptions:} For above properties of the
$\reach{t,u}$ relation to hold, we require that $\System$ be
deterministic. We also assume that $\System$ can be simulated from a
given initial state using function $\simulate^{\System}$. This
requires existence and uniqueness of trajectories over a finite time
horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
the vector field in each hybrid mode and ruling away issues such as
finite escape times~\cite{Meiss/2007/Differential}. From a practical
standpoint, these assumptions do not pose significant restrictions for
the type of problems that we wish to address.  Finally, we assume full
observability of the system state, ignore numerical errors due to
simulation and {\em do not} assume that an analytic closed-form
representation of the system dynamics is available.

@@

In statistics, linear regression is the problem of finding an affine
function or \textit{predictor}, which can `best' summarize the
relationship between the given set of observed input $\x$ and output
$\y$ vectors. The notion of `best' is formally captured using a loss
function which can be non-linear. We use the commonly used loss
function \textit{ordinary least squares} (OLS) for this presentation.

@@

The
Let the data set be comprised of $N$ input and output pairs $(\x, y)$,
where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is the case in
the current context of \textit{finding the best fit}, the problem is
over-determined; there are more equations than unknowns. Hence, a
single affine function cannot be found which satisfies the equation
$\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we need
to find the `best' choice for $\vec{a}$ and $b$. This is formally
defined using a loss function. For the case of simple linear
regression or OLS, the loss function is the sum of squares of the
errors in prediction.  The task is then to determine the vector of
coefficients $\vec{a}$ and an offset constant $b$, such that the least
square error of the affine predictor is minimized for the given data
set.
\begin{equation}
    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
\label{reg}
\end{equation}
To ease the presentation, we can rewrite the above as a homogeneous
expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
\[
    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
\]
The solution of OLS can be analytically computed as
\[ Ab = (X^TX)^{-1}X^T\y\]
where $X$ is the matrix representing the horizontal stacking of all
$\hat{\x}$.
The details can be found in several texts on learning and statistics
\cite{friedman2001elements}.

@@

% \mypara{Regression Analysis}

% In statistics, regression is the problem of finding a
% \textit{predictor}, which can suitably predict the relationship
% between the given set of observed input $\x$ and output $\y$ vectors.
% In other words, assuming that $\y$ depends on $\x$, regression
% strategies find either a parameterized or a non-parameterized
% prediction function to explain the dependence. We now discuss simple
% linear regression, which is parametric in nature and searches for an
% affine predictor. It is also called \textit{ordinary least squares}.

% \mypara{Ordinary Least Squares (OLS)}

% Let the data set be comprised of $N$ input and output pairs $(\x,
% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
% the case in the current context of \textit{finding the best fit}, the
% problem is over-determined; there are more equations than
% unknowns. Hence, a single affine function cannot be found which
% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
% defined using a loss function. For the case of simple linear
% regression or OLS, the loss function is the sum of squares of the
% errors in prediction.  The task is then to determine the vector of
% coefficients $\vec{a}$ and an offset constant $b$, such that the least
% square error of the affine predictor is minimized for the given data
% set.
% \begin{equation}
%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
% \label{reg}
% \end{equation}
% To ease the presentation, we can rewrite the above as a homogeneous
% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
% \[
%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
% \]
% The solution of OLS can be analytically computed as
% \[ Ab = (X^TX)^{-1}X^T\y\]
% where $X$ is the matrix representing the horizontal stacking of all
% $\hat{\x}$.
% The details can be found in several texts on learning and statistics
% \cite{friedman2001elements}.

% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
%     global relational model for the system. The data set for the
%     regression is a set of pairs $\setof{(\x,\x') | \x' =
%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
%     set can be generated to compute the error $\delta$ as an interval
%     of vectors, where each element is an interval $\delta_i \in
%     [\delta_{min}, \delta_{max}]$.
%
%     Affine maps are poor
%     approximations for arbitrary non-linear functions. Hence, we use a
%     collection of affine maps to approximate the local behaviors (in
%     state-space) of the system $\simulate$. This results in a piece
%     wise approximation, as described in the next section.

\subsection{Relational Abstractions}
For a hybrid automaton model, they can summarize the continuous
dynamics of each mode using a binary relation over continuous states.
The resulting relations are timeless (independent) and hence valid for
all time as long as the mode invariant is satisfied. The relations
take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
represents one of the relational operators $=, \ge, \le, <, >$. For
example, an abstraction that captures the monotonicity with respect to
time for the differential equation $\dot{x} = 2$ is $x' > x$. The
abstraction capturing the relation between the set of ODEs: $\dot{x} =
2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.

%%  Similar to timeless relational abstractions, time-aware relational
%%  abstractions \cite{mover2013time} construct binary relations between
%%  the current state $\x$ of the system and any future reachable state
%%  $\x'$. The difference lies with the latter also constructing relations
%%  between the current time $t$ and any future time $t'$. This is
%%  achieved by a case by case analysis of the eigen structure of the
%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
%%  case of linear systems with constant rate, real eigen values and
%%  complex eigenvalues.

@@

We can now use a PWA transition system $\rho$ to approximate the
behavior of a hybrid dynamical system $\System$ defined by a simulator
$\simulate^{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
form, we can only approximate general hybrid systems.}. Each
transition of $\rho$ represents a discrete time step $\Delta$. Let the
state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
predicates $g_\tau$ are given by a conjunction of hyper-planes and
hence can be represented in matrix form as a polyhedron (defined by
$m$ constraints) in the state-space
\[
g:C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
$m$. The affine maps $f_\tau$ can be represented in matrix form as
$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
an error term $\delta$ which is defined by a vector of intervals
$[\delta_{\min},\delta_{\max}]$. Hence,
\[
f:\x \mapsto A\x + \vb + \delta
\]
defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
\in f_\tau(\x)}$.

@@

\subsection{Learning Affine Function from Trajectory Segments}

The $\simulate$ function can be used to generate a data set of
relations parameterized by time $\Delta_t$, for different values of
states $\x_i$ and inputs $\vu_i$ as $\ds_{\Delta_t} =
\setof{(\x_i,\x'_i,\vu_i) | \x'_i = \simulate(\x_i,\vu_i)}$.
Moreover, by repeatedly applying $\simulate$, a discrete time series
of a finite length $N$ can be generated from a given state $\x_1$
and a sequence of inputs $\tupleof{\vu_1, \vu_2,\ldots,\vu_n}$ as shown
below.
\[\tupleof{\x_1, \x_2, \ldots, \x_n} \;\text{where}\; \x_{i>1} =
\simulate(\x_{i-1}, \vu_{i-1}, \Delta_t)\]

We now introduce a refinement of the data set $\ds \subseteq \ds$ by
incrementally observing the future states in the time series.

\[\ds_{\Delta_t} = \setof{(\x_i,\x'_i,\vu_i) | \x'_i =
\simulate(\x_i,\vu_i)}\]

\[\ds^{C_i} \subseteq \ds_t\]
\[\ds^{C_i}\]

$k$-Refinement of the data set $\ds^k$.
$\ds_k = \setof{(\x,\x',\vu) | \x' =  \simulate(\x,\vu,k\Delta_t)}$

state-input pairs of the form
$(\x,\x')$
data set equivalence traj segment
$D_c, $

### 2017-04-07 pwarel.tex
.

\subsection{Overview}
\label{sec:ovw}

@@

Instead of using a CEGAR like loop (as in \cite{zutshi2014multiple}), we use the
generated trajectory segments to learn quantitative models describing
the local behavior of the system. These models are defined by a set of
relations $R \subseteq \HybridStates \times \HybridStates$ for each
edge of the reachability graph.

>>>>

@@

Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
we use the generated trajectory segments to learn quantitative models
describing the local behavior of the system. These models are defined
by a set of relations $R \subseteq \HybridStates \times \HybridStates$
for each edge of the reachability graph.
